Idealized Wind Models of Cool Gas Outflows

### 2010-09-03 Winds/papers/Theory/First/Submit/ms.tex
blah
git-svn-id: https://svn.ucolick.org/xsvn/trunk/radtrans@5149 ac11df5e-a788-c23b-a8e1-c758c323924d
profx
git-svn-id: https://svn.ucolick.org/xsvn/trunk/radtrans@5174 ac11df5e-a788-c23b-a8e1-c758c323924d

\section{Introduction}

Nearly all gaseous objects that shine are also
observed to generate gaseous flows.  This includes the jets of protostars, the
stellar winds of massive O and B stars, the gentle Solar wind
of our Sun, the associated absorption of bright quasars, and the
spectacular jets of radio-loud AGN.   These gaseous outflows moderate
the accretion of new material onto the object, inject energy and
momentum into gas on large scales, and ...
Developing a comprehensive model for these flows is critical to
understanding the evolution of the source and its impact on the
surrounding environment.

Star-burst galaxies, whose luminosity is dominated by \ion{H}{2} regions
and massive stars, are also observed to drive gaseous outflows.  These
flows are generally expected (and sometimes observed) to have multiple
phases, for example a hot and diffuse phase traced by X-ray emission
together with a cool, denser phase traced by H$\alpha$ emission 
\citep[e.g.][]{shc+04,km10}. 
Several spectacular
examples in the local universe demonstrate that flows can extend to
tens of kpc from the galaxy \citep{M87} carrying significant speed
to escape from the gravitational potential well of the galaxy's dark
matter halo \citep{wind_escape}.

Beyond $z \sim 0$, galactic outflows are 
revealed by UV and optical absorption lines, e.g.\ \ion{Na}{1},
\ion{Mg}{2}, \ion{Si}{2} and \ion{C}{4} transitions.  With the galaxy
as a backlight, one observes gas that is predominantly
blue-shifted which indicates a flow toward
Earth and away from the galaxy.  These transitions are sensitive to
the cool (\ion{Mg}{2}; $T \sim 10^4$K) and warm (\ion{C}{4}; $T \sim
10^5$K) phases of the flow.  The incidence of cool gas outflows is
nearly universal in star-forming galaxies;  this includes systems at $z \sim 0$
which exhibit \ion{Na}{1} and \ion{Mg}{2} absorption
\citep{rvs05a,martin05,smn+09}, $z \sim 1$ star-forming galaxies traced by
\ion{Fe}{2} and \ion{Mg}{2} transitions \citep{wcp+09,rwk+10}, and
$z>2$ Lyman break galaxies (LBGs) that show blue-shifted \ion{Si}{2},
\ion{C}{2}, and \ion{O}{1} transitions \citep{sgp+96,lkg+97,shapley03}.

The observation of metal-line absorption is now a well established
means of identifying outflows. Very little research,
however,  has been directed
to comparing the observations against (even idealized) wind
models.  Instead, researchers have gleaned what limited information
is afforded by direct analysis of the absorption lines.  The data do yield
robust measurements of the speed of the gas yet they poorly constrain the
optical depth, covering fraction, density, temperature, and distance
of the flow from the galaxy.   In turn, constraints related to the
mass, energetics, and momentum of the flow suffer from
orders of magnitude uncertainty.  Both the origin and impact of
galactic-scale winds, therefore, remain open matters of debate
\citep{debate}.

Recent studies of $z \sim 1$ star-forming galaxies have revealed that
the cool gas often exhibits significant resonant-line emission (e.g.\
\ion{Mg}{2}) in
tandem with the nearly ubiquitous blue-shifted absorption
\citep{wcp+09,rwk+10}.  The resultant spectra resemble the P-Cygni
profile characteristic of stellar winds.
\cite{rubin+10c} have also detected line
emission from non-resonant \ion{Fe}{2}$^*$ transitions 
Their interpretation is that the emission is the result of photons
scattered off the wind into our sightline, analogous to P-Cygni.
\cite{rubin+10c} also observed spatially extended \ion{Mg}{2} emission
which they use to estimate the size of the outflow.  They infer that
the wind extends to at least 5\,kpc from the galaxy.  Line emission is
also observed for $z \sim 3$ LBGs in the resonant \lya\ transition
and non-resonant \ion{Si}{2}$^*$ transitions \citep{prs+02,shapley03}.
A comprehensive analysis of the emission lines
related to galactic-scale outflows
(e.g.\ via deep integral-field-unit observations) may offer unique
diagnostics on the morphology and density of the outflow, eventually
setting tighter constraints on the energetics of the flow.

[Phillips first reported?]
Bright \ion{Mg}{2} and \ion{Fe}{2}* line emission has also been
reported for individual galaxies at $z \sim 0.7$ by \cite{rubin+10c}
and Rubin et al.\ 2011 (in prep).
\cite{wcp+09}, who
studied \ion{Mg}{2} absorption in $z \sim 1.4$ galaxies, reported
\ion{Mg}{2} emission in a small subset of the individual galaxy
spectra of their large sample.  These were excluded from the full
analysis on concerns that the emission was related to AGN activity.
The stacked spectra of the remaining galaxies, however, also indicated
\ion{Mg}{2} emission, both directly and when the authors modeled and
`removed' the $v \approx 0\mkms$ absorption component.  Although the
authors suggested the emission could be related to back-scattered
light in the wind, they also allowed that it could be related to weak
AGN activity.   Very similar \ion{Mg}{2} emission was observed by
\cite{rwk+10} who repeated the analysis of \cite{wcp+09} on a set
of lower redshift galaxies. 
[Did Kate see FeII* emission in her stacks? If not, why not?]

Although astronomers are rapidly producing a wealth of observational
datasets on galactic-scale winds, a key ingredient to a proper
analysis is absent.
[Draw analogy to exploring SN lightcurves and spectra]

In this paper, we take the first steps toward modeling the absorption
and emission properties of cool gas outflows.  Using Monte Carlo
radiative transfer techniques, we study the nature of \ion{Mg}{2} and
\ion{Fe}{2} absorption and emission for winds with a range of
properties.  Although the winds are idealized, the results frequently
contradict our intuition and 
challenge the straightforward conversion of observables to (even crude) physical
constraints.  [Add another line or two]

The paper is organized as follows.  In $\S$~\ref{sec:method}, we
describe the methodology of our radiative transfer algorithms.  These
are applied to a fiducial wind model in $\S$~\ref{sec:fiducial} and
variations of this model in $\S$~\ref{sec:variants}.  In
$\S$~\ref{sec:alternate}, we explore wind models with a broader range
of density and velocity laws.  Consideration of key observables is
given in $\S$~\ref{sec:obs} and we discuss the principal results
in $\S$~\ref{sec:discuss}.  A brief summary is given in
$\S$~\ref{sec:summary}.

@@

Nearly all of the absorption studies of galactic-scale outflows 
have focused on intensely, star-forming galaxies.  The
intrinsic emission of these galaxies is a complex combination of
stars and \ion{H}{2} regions that is then modulated by dust and gas
within the interstellar medium (ISM).  For the spectral regions studied
here, most stars show a featureless continuum but a few spectral
types do show significant \ion{Mg}{2} and \ion{Fe}{2} absorption.
{\bf [Kate: Any emission??]}  \ion{H}{2} regions, meanwhile, are observed to emit
at the \mgiid\ doublet, primarily due to recombinations in the outer
layers {\bf [Kate: How about FeII?]}.  It is beyond the scope of this paper to
properly model the 
stellar absorption and \ion{H}{2} region emission, but the reader
should be aware that they can complicate the observed spectrum,
independent of the outflow, especially at velocities $v \approx 0 \mkms$.
In the following, we assume a simple flat continuum 
normalized to unit value.  The size of the emitting
region $r_{\rm source}$ is a free parameter, but we restrict its value
to be smaller than the minimum radial extent of any gaseous
component.   Lastly, the source does not absorb any scattered or
emitted photons.

\subsection{Monte Carlo Algorithm}

@@

For the majority of models studied in this paper, we assume the gas
contains no dust.
This is an invalid assumption, especially for material associated with
the ISM of a galaxy.  
Extraplanar material likely associated with a galactic-scale 
outflow has been observed to emit IR radiation characteristic of dust \citep[e.g.][]{M87_dust}.
Essentially all astrophysical environments that contain both cool gas
and metals also show signatures of dust depletion and extinction.  This includes the
ISM of star-forming and \ion{H}{1}-selected galaxies
\citep[e.g.][]{ss96,pw01,pcd+07}, strong \ion{Mg}{2} metal-line
absorption systems \citep{ykv+06,mnt+08}, and the galactic winds traced
by low-ion transitions \citep{prs+02,rvs05b}.  Although the galactic winds
traced specifically by \ion{Mg}{2} and \ion{Fe}{2} transitions have not (yet) been
demonstrated to contain dust, it is reasonable to consider its
effects.
In addition, \cite{msf+10} have argued from a statistical analysis
that dust is distributed to many tens kpc from $z \sim 0.1$ galaxies
and have suggested it was transported from the galaxies by
galactic-scale winds.  One is motivated, therefore, to consider
models that include the effects of dust extinction.

@@

In a few models, we include absorption by dust under the following
assumptions:
(i) the dust opacity scales with the density of the gas (i.e.\ we
adopt a fixed dust-to-gas ratio);
(ii) the opacity is independent of wavelength, a reasonable
approximation given the small spectral range analyzed;
(iii) the photons absorbed by dust are re-emitted at IR wavelengths
and are `lost' from the system.  The dust absorption is normalized by \taud, 
the integrated opacity of dust from the center of the system to
infinity.  The ambient ISM of a star-forming galaxy may be expected to exhibit
\taud\ values of one to a few at $\lambda \sim 3000$\AA\ \citep[e.g.][]{dust_again}.

@@

For a wind with a steep velocity law (i.e.\ a large gradient $dv/dr$)
and/or a a narrow intrinsic profile (i.e.\ a small Doppler parameter),
\citet{sobolev60} introduced a formalism that defines the opacity of
the wind at a given radius $\tau(r)$ in terms of the density and
velocity gradient of the flow at that radius;

@@

Focusing further on the \ion{Mg}{2} absorption, one notes that the profiles lie
well above zero intensity and have similar depth even though their $f\lambda$
values differ by a factor of two.  In standard absorption line
analysis, this is 
the tell-tale signature of a `cloud' that has a high optical depth (i.e.\
saturated) which only partially covers the emitting source
\citep[e.g.][]{hammann+10}.  Our fiducial wind model, however, 
{\it entirely covers the source}; the apparent partial covering must
be related to a different effect.
Figure~\ref{fig:noemiss} further emphasizes this point by comparing the 
absorption profiles from Figure~\ref{fig:fiducial_1d} against an
artificial model where no absorbed photons are 
re-emitted.   As expected from the
$\tau_{2796}$ profile (Figure~\ref{fig:fiducial_nvt}), this
`intrinsic' model
produces a strong \mgiid\ doublet that absorbs all photons at
$\delta v \approx -100 \mkms$, i.e.\ $I_{2796}^{\rm min} = \exp(-\tau^{\rm
  max}_{2796}) \approx 0$.
The true model, in contrast, has been `filled in' at $\delta v \approx -100
\mkms$ by photons scattered in the wind.  An
absorption-line analysis that ignores these effects
would (i) systematically underestimate the true optical
depth and (ii) falsely conclude that the wind partially covers the
source.  We will find that this
[behavior] is a generic result, even for wind models that include
dust and are not fully isotropic.

Turning to the emission profiles of the \mgiid\ doublet, one notes
that they are also similar with comparable equivalent widths The
flux of the \mgiib\ transition even exceeds that for \mgiia\ giving a
line ratio that is far below the $2:1$ ratio that one may have naively
expected. 
[Does recombination predict 2:1??]
The emission profiles are very similar because the gas is optically
thick yielding comparable total absorption. 
Furthermore, the flux from \mgiib\ actually exceeds the
\mgiia\ emission because the wind speed is greater than the velocity separation
of the doublet, $|\mvr|_{\rm max} > (\Delta v)_{\rm MgII}$.
Therefore, the red wing of the
\mgiia\ emission profile is partially absorbed by \mgiib\ and
re-emitted at lower frequency.  We conclude that the relative
strengths of the emission lines is sensitive to both the opacity and
velocity extent of the wind.

@@

The fiducial model assumes an
isotropic wind with only radial variations in velocity and density. 
Angular isotropy is obviously an idealized case, but
it is frequently assumed in studies of galactic-scale outflows
\citep[e.g.][]{steidel+10}.   There are several reasons, however, to
consider anisotropic winds.  Firstly, galaxies are not spherically
symmetric;  the sources driving the
wind (e.g.\ supernovae, AGN) are very unlikely to be isotropically distributed
within the galaxy.  
Secondly, the galactic ISM frequently has a disk-like morphology
which will suppress the wind preferentially at low galactic latitudes,
perhaps yielding a bi-conic morphology \citep[e.g.][]{M87}.
Lastly, the galaxy could be surrounded by an
anisotropic gaseous halo whose interaction would produce an irregular 
outflow.

With these considerations in mind, we reanalyzed the fiducial model
with the 3D algorithm after departing from isotropy.  It is beyond the
scope of this paper to explore a full suite of anisotropic profiles.
We consider two simple examples: (i) half the fiducial model, where 
the density is set to zero for $2\pi$ steradians.  This model is
viewed from $\phi = 0^\circ$ (source uncovered) 
to $\phi = 180^\circ$ (source covered); and 
(ii) a bi-conical wind where the density is set to zero for angles
$|\theta| < \theta_b$ that and is viewed along the plane of symmetry.

@@

In Figure~\ref{fig:dust}, we show the \ion{Mg}{2} and \ion{Fe}{2}
profiles of the fiducial model ($\mtaud = 0$) against a series of
models with $\mtaud > 0$.  For the \ion{Mg}{2} transitions, the
dominant effect is the suppression of line emission at $\mdv \ge 0
\mkms$.  These `red' photons have scattered off the
backside of the wind and must travel a longer path than other
photons.  Dust leads to a differential reddening that increases with 
velocity relative to line-center. This is a natural consequence of dust
extinction and is most evident in the \feiic\ 
emission profile which is symmetrically distributed around
$\mdv= 0 \mkms$ in the $\mtaud=0$ model.   
The degree of suppression of the line-emission is relatively modest,
however.  Specifically, we find that the flux is reduced by a factor 
of the order of (1+\taud)$^{-1}$ (Figure~\ref{fig:dust_tau},
instead of the factor
exp(--\taud) that one may have naively predicted. 
Although the \ion{Mg}{2} photons are resonantly
In terms of absorption, the profiles are
nearly identical for $\mtaud \le 3$.  One requires very high
extinction to eventually produce a deepening of the profiles at 
$\mdv \approx -100 \mkms$.

@@

[Kate: Write the details of velocity + density law]

\begin{equation}
\mvr(r) = 2\sigma \sqrt{R_g \ltp \frac{1}{R_0} - \frac{1}{r} \rtp
   + \ln\ltp R_0/r \rtp }
\end{equation}

\begin{equation}
n(r) = \frac{dM_{\rm wind}/dt}{r^2 v(r)}
\end{equation}
Similar to the fiducial wind model, this wind has a decreasing density
law with $n_{\rm H}$ roughly proportional to $r^{-2}$. 
In contrast to the fiducial model, the velocity law is nearly
constant  with radius before decreasing sharply as r approaches $R_0$.
To produce a wind whose \mgiia\ optical depth profile peaks at 
$\tau_{2796}^{\rm max} \gtrsim 10$, we set $R_g = r_{\rm inner} =
1\,\rm kpc$, $R_0=20\,\rm kpc$, and $\sigma = 125 \mkms$.  The density
and velocity peak at $r=r_{\rm inner}$ where $n_{\rm H}^0 = 0.07
\cm{-3}$ and $\mvr = 380 \mkms$.  This corresponds to a mass flow of
$dM_{\rm wind}/dt = 0.63 \msun\; \rm yr^{-1}$.

>>>>

\section{Introduction}
\label{sec:intro}

Nearly all gaseous objects that shine are also
observed to generate gaseous flows.  This includes the jets of protostars, the
stellar winds of massive O and B stars, the gentle Solar wind
of our Sun, the associated absorption of bright quasars, and the
spectacular jets of radio-loud AGN.   These gaseous outflows 
%%KHRR
regulate the metal and dust content and distribution within the 
objects and its surroundings, 
%%
moderate the accretion of new material, and %onto the object, 
inject energy and momentum into gas on large scales. %, and ...
Developing a comprehensive model for these flows is critical to
understanding the evolution of the source and its impact on the
surrounding environment.

Starburst galaxies, whose luminosity is dominated by \ion{H}{2} regions
and massive stars, are also observed to drive gaseous outflows.  These
flows are generally expected (and sometimes observed) to have multiple
phases, for example a hot and diffuse phase traced by X-ray emission
together with a cool, denser phase traced by H$\alpha$ emission 
%% KHRR -- added refs
\citep[e.g.][]{ham90,martin99,shc+04,km10}. 
Several spectacular
examples in the local universe demonstrate that flows can extend to
up to $\sim 10$ kpc
%tens of kpc 
from the galaxy \citep{M87,lhw99,wsg08,vsr+03} carrying significant speed
to escape from the gravitational potential well of the galaxy's dark
matter halo \citep{wind_escape,sh09}.

%Beyond $z \sim 0$, 
Galactic outflows are also
revealed by UV and optical absorption lines, e.g.\ \ion{Na}{1},
\ion{Mg}{2}, \ion{Si}{2} and \ion{C}{4} transitions.  With the galaxy
as a backlight, one observes gas that is predominantly
blue-shifted which indicates a flow toward
Earth and away from the galaxy.  These transitions are sensitive to
the cool (\ion{Mg}{2}; $T \sim 10^4$K) and warm (\ion{C}{4}; $T \sim
10^5$K) phases of the flow.  
The incidence of cool gas outflows is
nearly universal in 
%% KHRR added:
vigorously
%%
star-forming galaxies;  this includes systems at $z \sim 0$
which exhibit \ion{Na}{1} and \ion{Mg}{2} absorption
\citep{rvs05a,martin05,smn+09,mb09,cth+10}, 
$z \sim 1$ star-forming galaxies traced by
\ion{Fe}{2} and \ion{Mg}{2} transitions \citep{wcp+09,rwk+10}, and
$z>2$ Lyman break galaxies (LBGs) that show blue-shifted \ion{Si}{2},
\ion{C}{2}, and \ion{O}{1} transitions \citep{sgp+96,lkg+97,shapley03}.

The observation of metal-line absorption is now a well-established
means of identifying outflows.  Furthermore, because the X-ray and
H$\alpha$ emission generated by winds is faint, absorption-line
analyses have traditionally been the only way to probe outflows in
distant galaxies. However, very little research
has been directed
toward comparing the observations against (even idealized) wind
models (although see \citealt{fmm+09}).  
Instead, researchers have gleaned what limited information
is afforded by direct analysis of the absorption lines.  The data do yield
robust measurements of the speed of the gas (is this really true?), 
yet they poorly constrain the
optical depth, covering fraction, density, temperature, and distance
of the flow from the galaxy.   In turn, constraints related to the
mass, energetics, and momentum of the flow suffer from
orders of magnitude uncertainty.  Both the origin and impact of
galactic-scale winds, therefore, remain open matters of debate
\citep{debate,mqt05,sdr08,sh09,ssr02,od06,kkd+09}.

Recent studies of $z \lesssim 1$ star-forming galaxies have revealed that
the cool outflowing gas often exhibits significant resonant-line emission (e.g.\
\ion{Mg}{2}, \ion{Na}{1}) in
tandem with the nearly ubiquitous blue-shifted absorption
\citep{wcp+09,rwk+10,cth+10}.  The resultant spectra resemble the P-Cygni
profile characteristic of stellar winds.
This phenomenon was first reported by \citet{phillips93}, who observed 
blue-shifted absorption and red-shifted emission in the \ion{Na}{1}
transition in the spectrum of the local starburst galaxy NGC 1808.
More recently, \cite{wcp+09}, who
studied \ion{Mg}{2} absorption in $z \sim 1.4$ galaxies, reported
\ion{Mg}{2} emission in a small subset of the individual galaxy
spectra of their large sample.  These were excluded from the full
analysis on concerns that the emission was related to AGN activity.
The stacked spectra of the remaining galaxies, however, also indicated
\ion{Mg}{2} emission, both directly and when the authors modeled and
`removed' the $v \approx 0\mkms$ absorption component.  Although the
authors suggested the emission could be related to back-scattered
light in the wind, they also allowed that it could be related to weak
AGN activity.   Very similar \ion{Mg}{2} emission was observed by
\cite{rwk+10} who repeated the analysis of \cite{wcp+09} on a set
of lower redshift galaxies.

Bright \ion{Mg}{2} line emission has also been
reported for individual galaxies at $z \sim 0.7$ by \cite{rubin+10c}
and Rubin et al.\ 2011 (in prep).
In their analysis of a single galaxy spectrum, \cite{rubin+10c}
further demonstrated that the \ion{Mg}{2} line emission is spatially
extended, and used the size of the emission to infer that the wind 
extends to at least 7\,kpc from the galaxy.
These authors additionally detected line
emission from non-resonant \ion{Fe}{2}$^*$ transitions, and attribute
the emission to fluorescence powered by \ion{Fe}{2} resonant
absorption.  In other words, these photons are re-emitted by the wind
into our sightline, and are analogous to the emitted photons in a P
Cygni profile.
%Their interpretation is that the emission is the result of photons
%scattered off the wind into our sightline, analogous to P-Cygni.
%\cite{rubin+10c} also observed spatially extended \ion{Mg}{2} emission
%which they use to estimate the size of the outflow.  They infer that
%the wind extends to at least 7\,kpc from the galaxy.  
Line emission is
also observed for $z \sim 3$ LBGs in the resonant \lya\ transition
and non-resonant \ion{Si}{2}$^*$ transitions \citep{prs+02,shapley03}.
We emphasize that this emission arises from a completely different
physical process than those generating X-ray and H$\alpha$ emission
(e.g., shocks), and presumably probes both the cool gas at the base of
the wind and the outskirts of the flow (i.e., wherever a given
transition is optically thick).
A comprehensive analysis of the %emission lines
scattered and fluorescent emission
related to galactic-scale outflows
(e.g.\ via deep integral-field-unit [IFU] observations,
\citealt{ssb+05,wsg08}) 
may offer unique
diagnostics on the spatial extent, morphology, and density of the outflow 
from distant galaxies,
eventually
setting tighter constraints on the energetics of the flow.

%[Phillips first reported?]
%Bright \ion{Mg}{2} and \ion{Fe}{2}* line emission has also been
%reported for individual galaxies at $z \sim 0.7$ by \cite{rubin+10c}
%and Rubin et al.\ 2011 (in prep).
%\cite{wcp+09}, who
%studied \ion{Mg}{2} absorption in $z \sim 1.4$ galaxies, reported
%\ion{Mg}{2} emission in a small subset of the individual galaxy
%spectra of their large sample.  These were excluded from the full
%analysis on concerns that the emission was related to AGN activity.
%The stacked spectra of the remaining galaxies, however, also indicated
%\ion{Mg}{2} emission, both directly and when the authors modeled and
%`removed' the $v \approx 0\mkms$ absorption component.  Although the
%authors suggested the emission could be related to back-scattered
%light in the wind, they also allowed that it could be related to weak
%AGN activity.   Very similar \ion{Mg}{2} emission was observed by
%\cite{rwk+10} who repeated the analysis of \cite{wcp+09} on a set
%of lower redshift galaxies. 
%[Did Kate see FeII* emission in her stacks? If not, why not?]

Although astronomers are rapidly producing a wealth of observational
datasets on galactic-scale winds, a key ingredient to a proper
analysis is absent.  Just as comparisons between observed supernova 
lightcurves and spectra and radiative transfer calculations of
of SNe models have provided crucial insight into the physics driving,
e.g., standard candle relationships for both Type Ia and IIP SNe 
\citep{kw07,kw09}, modeling of the observable signatures of galactic
outflows is necessary for understanding the physical
properties of the gas and ultimately the physics driving the flows. 
%[Draw analogy to exploring SN lightcurves and spectra]
In this paper, we take the first steps toward modeling the absorption
and emission properties of cool gas outflows as observed in one
dimensional spectra and by IFUs.  Using Monte Carlo
radiative transfer techniques, we study the nature of \ion{Mg}{2} and
\ion{Fe}{2} absorption and emission for winds with a range of
properties, accounting for the effects of resonant scattering and
fluorescence for the first time.  
Although the winds are idealized, the results frequently
contradict our intuition and 
challenge the straightforward conversion of observables to (even crude) physical
constraints.  We caution that these findings have a direct bearing on
recent and upcoming surveys of galactic outflows, particularly those
which make use of \ion{Mg}{2} and \ion{Fe}{2} transitions to probe
outflow properties.
%[Add another line or two]

The paper is organized as follows.  In $\S$~\ref{sec:method}, we
describe the methodology of our radiative transfer algorithms.  These
are applied to a fiducial wind model in $\S$~\ref{sec:fiducial} and
variations of this model in $\S$~\ref{sec:variants}.  In
$\S$~\ref{sec:alternate}, we explore wind models with a broader range
of density and velocity laws. We discuss the principal results and
connect to observations in $\S$~\ref{sec:discuss}.  A brief summary is
given in $\S$~\ref{sec:summary}.

@@

Nearly all of the absorption studies of galactic-scale outflows 
have focused on intensely, star-forming galaxies.  The
intrinsic emission of these galaxies is a complex combination of
stars and \ion{H}{2} regions that is then modulated by dust and gas
within the interstellar medium (ISM).  For the spectral regions studied
here, most stars show a featureless continuum, but a few spectral
types do show significant \ion{Mg}{2} and \ion{Fe}{2} absorption.
In addition, asymmetric and/or blue-shifted absorption is exhibited in these
transitions in A and F stars driving stellar winds \citep{sll+94}.
\ion{Mg}{2} P Cygni profiles are observed in a handful of F stars by \cite{sll+94}, who
attribute the emission to chromospheric activity rather than mass-loss effects.
%{\bf [Kate: Any emission??]}  
\ion{H}{2} regions, meanwhile, are observed to emit weakly
at the \mgiid\ doublet, primarily due to recombinations in the outer
layers \citep{kbc+93} {\bf [Kate: How about FeII?  X: I don't know of any examples of this.]}.  
It is beyond the scope of this paper to
properly model the 
stellar absorption and \ion{H}{2} region emission, but the reader
should be aware that they can complicate the observed spectrum,
independent of the outflow, especially at velocities $v \approx 0 \mkms$.
In the following, we assume a simple flat continuum 
normalized to unit value.  The size of the emitting
region $r_{\rm source}$ is a free parameter, but we restrict its value
to be smaller than the minimum radial extent of any gaseous
component.   Lastly, the source does not absorb any scattered or
emitted photons.

\subsection{Monte Carlo Algorithm}
\label{sec:monte}

@@

For the majority of models studied in this paper, we assume the gas
contains no dust.
This is an invalid assumption, especially for material associated with
the ISM of a galaxy.  
Extraplanar material likely associated with a galactic-scale 
outflow has been observed to emit IR radiation characteristic of dust 
\citep[e.g.][]{M87_dust,hrg90,adb99,rkl01}.
Essentially all astrophysical environments that contain both cool gas
and metals also show signatures of dust depletion and extinction.  This includes the
ISM of star-forming and \ion{H}{1}-selected galaxies
\citep[e.g.][]{ss96,pw01,pcd+07}, strong \ion{Mg}{2} metal-line
absorption systems \citep{ykv+06,mnt+08}, and the galactic winds traced
by low-ion transitions \citep{prs+02,rvs05b}.  Although the galactic winds
traced specifically by \ion{Mg}{2} and \ion{Fe}{2} transitions have not (yet) been
demonstrated to contain dust, it is reasonable to consider its
effects.
In addition, \cite{msf+10} have argued from a statistical analysis
that dust is distributed to many tens kpc from $z \sim 0.1$ galaxies
and have suggested it was transported from the galaxies by
galactic-scale winds.  One is motivated, therefore, to consider
models that include the effects of dust extinction.

@@

In a few models, we include absorption by dust under the following
assumptions:
(i) the dust opacity scales with the density of the gas (i.e.\ we
adopt a fixed dust-to-gas ratio);
(ii) the opacity is independent of wavelength, a reasonable
approximation given the small spectral range analyzed;
(iii) the photons absorbed by dust are re-emitted at IR wavelengths
and are `lost' from the system.  The dust absorption is normalized by \taud, 
the integrated opacity of dust from the center of the system to
infinity.  The ambient ISM of a star-forming galaxy may be expected to exhibit
\taud\ values of one to a few at $\lambda \sim 3000$\AA\ \citep[e.g.][]{dust_again,cf00}.

@@

For a wind with a steep velocity law (i.e.\ a large gradient $dv/dr$)
and/or a narrow intrinsic profile (i.e.\ a small Doppler parameter),
\citet{sobolev60} introduced a formalism that defines the opacity of
the wind at a given radius $\tau(r)$ in terms of the density and
velocity gradient of the flow at that radius;

@@

Focusing further on the \ion{Mg}{2} absorption, one notes that the profiles lie
well above zero intensity and have similar depth even though their $f\lambda$
values differ by a factor of two.  In standard absorption line
analysis, this is 
the tell-tale signature of a `cloud' that has a high optical depth (i.e.\
saturated) which only partially covers the emitting source
\citep[e.g.][]{rvs05a,hammann+10}.  Our fiducial wind model, however, 
{\it entirely covers the source}; the apparent partial covering must
be related to a different effect.
Figure~\ref{fig:noemiss} further emphasizes this point by comparing the 
absorption profiles from Figure~\ref{fig:fiducial_1d} against an
artificial model where no absorbed photons are 
re-emitted.   As expected from the
$\tau_{2796}$ profile (Figure~\ref{fig:fiducial_nvt}), this
`intrinsic' model
produces a strong \mgiid\ doublet that absorbs all photons at
$\delta v \approx -100 \mkms$, i.e.\ $I_{2796}^{\rm min} = \exp(-\tau^{\rm
  max}_{2796}) \approx 0$.
The true model, in contrast, has been `filled in' at $\delta v \approx -100
\mkms$ by photons scattered in the wind.  An
absorption-line analysis that ignores these effects
would (i) systematically underestimate the true optical
depth and (ii) falsely conclude that the wind partially covers the
source.  We will find that this
[behavior] is a generic result, even for wind models that include
dust and are not fully isotropic.

%%% KHRR edited this-- no content change intended, just tried to
%%% remove repetition
Turning to the emission profiles of the \mgiid\ doublet, one notes
that they are quite similar with comparable equivalent widths.
This is because the gas is optically
thick yielding comparable total absorption. The
flux of the \mgiib\ transition even exceeds that for \mgiia\ 
because the wind speed is greater than the velocity separation
of the doublet, $|\mvr|_{\rm max} > (\Delta v)_{\rm MgII}$.
Therefore, the red wing of the
\mgiia\ emission profile is partially absorbed by \mgiib\ and
re-emitted at lower frequency.
This yields a
line ratio that is far below the $2:1$ ratio that one may have naively
expected, and leads us to conclude that the relative
strengths of the emission lines is sensitive to both the opacity and
velocity extent of the wind. 
[Does recombination predict 2:1??  The simple HII region Cloudy model
produces nearly a 2:1 ratio, though not exactly. (Kate)]
%The emission profiles are very similar because the gas is optically
%thick yielding comparable total absorption. 
%Furthermore, the flux from \mgiib\ actually exceeds the
%\mgiia\ emission because the wind speed is greater than the velocity separation
%of the doublet, $|\mvr|_{\rm max} > (\Delta v)_{\rm MgII}$.
%Therefore, the red wing of the
%\mgiia\ emission profile is partially absorbed by \mgiib\ and
%re-emitted at lower frequency.  We conclude that the relative
%strengths of the emission lines is sensitive to both the opacity and
%velocity extent of the wind.

@@

The fiducial model assumes an
isotropic wind with only radial variations in velocity and density. 
Angular isotropy is obviously an idealized case, but
it is frequently assumed in studies of galactic-scale outflows
\citep[e.g.][]{steidel+10}.   There are several reasons, however, to
consider anisotropic winds.  Firstly, galaxies are not spherically
symmetric;  the sources driving the
wind (e.g.\ supernovae, AGN) are very unlikely to be isotropically distributed
within the galaxy.  
Secondly, the galactic ISM frequently has a disk-like morphology
which will suppress the wind preferentially at low galactic latitudes,
perhaps yielding a bi-conic morphology \citep[e.g.][]{ham90,wws02}.
Lastly, the galaxy could be surrounded by an
anisotropic gaseous halo whose interaction would produce an irregular 
outflow.

With these considerations in mind, we reanalyzed the fiducial model
with the 3D algorithm after departing from isotropy.  It is beyond the
scope of this paper to explore a full suite of anisotropic profiles.
We consider two simple examples: (i) half the fiducial model, where 
the density is set to zero for $2\pi$ steradians.  This model is
viewed from $\phi = 0^\circ$ (source uncovered) 
to $\phi = 180^\circ$ (source covered); and 
(ii) a bi-conical wind where the density is set to zero for angles
$|\theta| < \theta_b$ and that is viewed along the plane of symmetry.

@@

In Figure~\ref{fig:dust}, we show the \ion{Mg}{2} and \ion{Fe}{2}
profiles of the fiducial model ($\mtaud = 0$) against a series of
models with $\mtaud > 0$.  For the \ion{Mg}{2} transitions, the
dominant effect is the suppression of line emission at $\mdv \ge 0
\mkms$.  These `red' photons have scattered off the
backside of the wind and must travel a longer path than other
photons.  Dust leads to a differential reddening that increases with 
velocity relative to line-center. This is a natural consequence of dust
extinction and is most evident in the \feiic\ 
emission profile which is symmetrically distributed around
$\mdv= 0 \mkms$ in the $\mtaud=0$ model.   
The degree of suppression of the line-emission is relatively modest,
however.  Specifically, we find that the flux is reduced by a factor 
of the order of (1+\taud)$^{-1}$ (Figure~\ref{fig:dust_tau}),
instead of the factor
exp(--\taud) that one may have naively predicted. 
%Although the \ion{Mg}{2} photons are resonantly
In terms of absorption, the profiles are
nearly identical for $\mtaud \le 3$.  One requires very high
extinction to eventually produce a deepening of the profiles at 
$\mdv \approx -100 \mkms$.

@@

%[Kate: Write the details of velocity + density law]
As discussed in $\S$~\ref{sec:intro}, the importance of various physical mechanisms
in driving these ubiquitous galactic outflows remains an open
question.  Several studies advocate a primarily energy-driven
wind scenario, in which the hot wind produced by the thermalized
energy from SNe ejecta entrains cool clouds via ram pressure \citep[e.g.][]{cc85,ham90,sh09}.
Alternatively, momentum deposition by photons emitted by luminous star clusters onto
dust grains in the ISM may also play an integral role in driving
large-scale outflows \citep[e.g.][]{mqt05,mmt10}.  Both of these
mechanisms likely contribute more or less significantly in a given
galaxy; however, the latter (radiation-pressure driving) has the added
advantage that it does not destroy cool gas clouds as it acts.  Here
we explore the line profiles produced by a radiation-pressure driven
wind model as described by \cite{mqt05}.
